The Journey from Wild to Textbook Data: A Case Study from the National Longitudinal Survey of Youth

### 2022-02-03 paper/paper.Rmd
censor yowie

For the field of statistics, and data science by association, it is increasingly important to reach the individual. One might describe this as a divergence of purpose, statistics for public policy or statistics for the public. The two are not the same. As the world becomes more electronically connected combating misinformation and mitigating conspiracy theories require that statistics address the individual. For example, with the wages data, the message for the individual is that you are more than your demographic. The majority of people tracked are not like the average, the average is an anomaly. If you have a bad experience, that your wages have declined over time, you are not alone, there are others like you, and more than you think.

@@

Our process of cleaning builds heavily on the `tidyverse` approach [@tidyverse]. The data is first organised into "tidy data" [@WickhamHadley2014TD] and then further wrangled using the data pipeline and split-apply-combine approach [@plyr]. <!--AE comment: this is an important point that shouldn't be relegated to a parenthetical statement. --> (Tidy data shouldn't be confused with "tame data" which @tamedata coined to refer to textbook data sets suitable for teaching, particularly teaching statistics.) The resulting (tame) data is provided in a new R package called `yowie` which includes the code so that the process is reproducible, and could be used to further refresh the data as new records are made available in the NLSY79 database.

>>>>

For the field of statistics, and data science by association, it is increasingly important to reach the individual. One might describe this as a divergence of purpose, statistics for public policy or statistics for the public. The two are not the same. As the world becomes more electronically connected combating misinformation and mitigating conspiracy theories require that statistics address the individual. For example, with the wages data, the message for the individual is that you are more than your demographic. The majority of people tracked are not like the average, the average is an anomaly. If you have a bad experience, that your wages have declined over time, you are not alone, there are others like you, and more than you think. [ET: a related article: https://www.smh.com.au/national/by-the-numbers--the-average-australian-doesnt-exist--not-a-single-one-of-us-is-normal-20150124-12xlh0.html]

@@

Our process of cleaning builds heavily on the `tidyverse` approach [@tidyverse]. The data is first organised into "tidy data" [@WickhamHadley2014TD] and then further wrangled using the data pipeline and split-apply-combine approach [@plyr]. <!--AE comment: this is an important point that shouldn't be relegated to a parenthetical statement. --> (Tidy data shouldn't be confused with "tame data" which @tamedata coined to refer to textbook data sets suitable for teaching, particularly teaching statistics.) The resulting (tame) data is provided in a new R package called `r yowie` which includes the code so that the process is reproducible, and could be used to further refresh the data as new records are made available in the NLSY79 database.
